---
title: "PlateCountAnalysis"
author: "Kevin Watanabe-Smith"
date: "June 16, 2016"
output: html_document
---
This is an attempt at analyzing the limiting dilution plating experiments in my
ongoing Acquired Mutations project. No analysis has been performed to this point.


```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(stringr)

RawData <- read.csv("PlateSortResults.csv")
summary(RawData)
```

We should start by getting an idea for Plating Efficiency (PE) from the WEHI wells

```{r}
WEHIonly <- RawData %>% filter(Media == "WEHI")
summary(WEHIonly)
WEHIcomplete <- WEHIonly %>% filter(!is.na(WellsGrown))
Wpercents <- WEHIcomplete %>% transform(GrowthPercent = WellsGrown / WellsPlated)
hist(Wpercents$GrowthPercent)
```

Which conditions gave the low end of the growth percent?

```{r}
Wlowgrowth <- Wpercents %>% filter(GrowthPercent < 0.4)
Wlowgrowth[1,c(1,5,6,9,10)]
```

Comparing PE of cells plated at 2 days and 9 days post-infection

```{r}
Wpercents2 <- Wpercents %>% filter(DaysPostInfection == 2)
Wpercents9 <- Wpercents %>% filter(DaysPostInfection == 9)
summary(Wpercents2)
summary(Wpercents9)
hist(Wpercents2$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 2 days post-infection", breaks = 8, main = "WEHI 2 days post infection")
curve(dnorm(x, mean = mean(Wpercents2$GrowthPercent), sd = sd(Wpercents2$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)
hist(Wpercents9$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 9 days post-infection", breaks = 8, main = "WEHI 9 days post infection")
curve(dnorm(x, mean = mean(Wpercents9$GrowthPercent), sd = sd(Wpercents9$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)

```

### **Conclusion: days post-infection do not seem to impact plating efficiency**

Let's try the same approach with the R10 data

```{r}
AllPercents <- RawData %>% mutate(GrowthPercent = WellsGrown / WellsPlated)
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 1, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 1 cell per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 100, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 100 cells per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 1000, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 1,000 cells per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 10000, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 10,000 cells per well")

```

For starters, we need to accomplish three points:
1. We need to adjust WellsPlated for Censored
2. We need to calculate GrowthPercent for all observations
3. We need to adjust GrowthPercent for Plating Effiency
```{r}
CompleteData <- RawData %>% filter(!is.na(WellsPlated))
CompleteData <- CompleteData %>% mutate(TrueWellsPlated = WellsPlated - WellsCensored)
CompleteData <- CompleteData %>% select(-WellsPlated, -WellsCensored)
head(CompleteData)

CompleteData <- CompleteData %>% mutate(LineID = str_sub(LineID, 2)) %>% mutate(LineID = paste(LineNum_legacy, LineID, sep = ""))
head(CompleteData)

CompletePercent <- CompleteData %>% mutate(GrowthPercent = WellsGrown / TrueWellsPlated) %>% select(-WellsGrown, -TrueWellsPlated, -LineNum_legacy, -Transgene, -Mutation)
head(CompletePercent)

```

Trying to reshape data to allow for better analysis

```{r}
#Change Media type to a single letter
CompletePercent <- CompletePercent %>% mutate(Media = str_sub(Media,1,1))
#Concatenate some identifiers into a single field, then drop unused columns
CompleteConcat <- CompletePercent %>% mutate(Media_Day_Cells = paste(Media, DaysPostInfection, CellsPerWell, sep = "_")) %>% select(-PlateSeries, -DaysPostInfection, -CellsPerWell, -Media)
#Change the identifier column into a factor (instead of character)
CompleteConcat <- CompleteConcat %>% mutate(Media_Day_Cells = as.factor(Media_Day_Cells))
head(CompleteConcat)

#Recast the data with LineID as the rows, Media_Day_Cells as the columns, and GrowthPercent as the values
CompleteTidy <- recast(CompleteConcat, LineID ~ Media_Day_Cells, id.var = c("LineID","Media_Day_Cells"), fun.aggregate = mean, na.rm = TRUE)
head(CompleteTidy)
#Output the results
write.csv(CompleteTidy, file = "CompleteTidy.csv")
```

Going to do some scatter plots to look for covariance between PE on day2 and day9 samples

```{r}
plot(CompleteTidy$`W_2_1`, CompleteTidy$`W_9_1`, xlab = "PE 2 Days post infection", ylab = "PE 9 Days post infection", xlim = 0:1, ylim = 0:1)
abline(a = 0, b = 1, col = "darkblue")
```

The PE of cell lines cluster, but not super related (IMO) which is expected.
## Adjusting Growth Percentage for Plating Efficiency
```{r}
summary(CompleteTidy)
CompleteTidy <- CompleteTidy %>% mutate(R_2_1 = R_2_1 / W_2_1)
summary(CompleteTidy)
```

