---
title: "PlateCountAnalysis"
author: "Kevin Watanabe-Smith"
date: "June 16, 2016"
output: html_document
---
This is an attempt at analyzing the limiting dilution plating experiments in my
ongoing Acquired Mutations project. No analysis has been performed to this point.


```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(stringr)

RawData <- read.csv("PlateSortResults.csv")
summary(RawData)
```

We should start by getting an idea for Plating Efficiency (PE) from the WEHI wells

```{r}
WEHIonly <- RawData %>% filter(Media == "WEHI")
summary(WEHIonly)
WEHIcomplete <- WEHIonly %>% filter(!is.na(WellsGrown))
Wpercents <- WEHIcomplete %>% transform(GrowthPercent = WellsGrown / WellsPlated)
hist(Wpercents$GrowthPercent)
```

Which conditions gave the low end of the growth percent?

```{r, cache = TRUE}
Wlowgrowth <- Wpercents %>% filter(GrowthPercent < 0.4)
Wlowgrowth[1,c(1,5,6,9,10)]
```

Comparing PE of cells plated at 2 days and 9 days post-infection

```{r, cache = TRUE}
Wpercents2 <- Wpercents %>% filter(DaysPostInfection == 2)
Wpercents9 <- Wpercents %>% filter(DaysPostInfection == 9)
summary(Wpercents2)
summary(Wpercents9)
hist(Wpercents2$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 2 days post-infection", breaks = 8, main = "WEHI 2 days post infection")
curve(dnorm(x, mean = mean(Wpercents2$GrowthPercent), sd = sd(Wpercents2$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)
hist(Wpercents9$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 9 days post-infection", breaks = 8, main = "WEHI 9 days post infection")
curve(dnorm(x, mean = mean(Wpercents9$GrowthPercent), sd = sd(Wpercents9$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)

```

### **Conclusion: days post-infection do not seem to impact plating efficiency**

Let's try the same approach with the R10 data

```{r, cache = TRUE}
AllPercents <- RawData %>% mutate(GrowthPercent = WellsGrown / WellsPlated)
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 1, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 1 cell per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 100, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 100 cells per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 1000, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 1,000 cells per well")
hist((AllPercents %>% filter(Media == "R10", CellsPerWell == 10000, !is.na(WellsGrown)))$GrowthPercent, main = "All growth percentages in R10, 10,000 cells per well")

```

For starters, we need to accomplish three points:
1. We need to adjust WellsPlated for Censored
2. We need to calculate GrowthPercent for all observations
3. We need to adjust GrowthPercent for Plating Effiency
```{r, cache = TRUE}
CompleteData <- RawData %>% filter(!is.na(WellsPlated))
CompleteData <- CompleteData %>% mutate(TrueWellsPlated = WellsPlated - WellsCensored)
CompleteData <- CompleteData %>% select(-WellsPlated, -WellsCensored)
head(CompleteData)

CompleteData <- CompleteData %>% mutate(LineID = str_sub(LineID, 2)) %>% mutate(LineID = paste(LineNum_legacy, LineID, sep = ""))
head(CompleteData)

CompletePercent <- CompleteData %>% mutate(GrowthPercent = WellsGrown / TrueWellsPlated) %>% select(-WellsGrown, -TrueWellsPlated, -LineNum_legacy, -Transgene, -Mutation)
head(CompletePercent)

```

Trying to reshape data to allow for better analysis

```{r, cache = TRUE}
#Change Media type to a single letter
CompletePercent <- CompletePercent %>% mutate(Media = str_sub(Media,1,1))
#Concatenate some identifiers into a single field, then drop unused columns
CompleteConcat <- CompletePercent %>% mutate(Media_Day_Cells = paste(Media, DaysPostInfection, CellsPerWell, sep = "_")) %>% select(-PlateSeries, -DaysPostInfection, -CellsPerWell, -Media)
#Change the identifier column into a factor (instead of character)
CompleteConcat <- CompleteConcat %>% mutate(Media_Day_Cells = as.factor(Media_Day_Cells))
head(CompleteConcat)

#Recast the data with LineID as the rows, Media_Day_Cells as the columns, and GrowthPercent as the values
CompleteTidy <- recast(CompleteConcat, LineID ~ Media_Day_Cells, id.var = c("LineID","Media_Day_Cells"), fun.aggregate = mean, na.rm = TRUE)
head(CompleteTidy)
#Output the results
#write.csv(CompleteTidy, file = "CompleteTidy.csv")
```

Going to do some scatter plots to look for covariance between PE on day2 and day9 samples

```{r, cache = TRUE}
plot(CompleteTidy$`W_2_1`, CompleteTidy$`W_9_1`, xlab = "PE 2 Days post infection", ylab = "PE 9 Days post infection", xlim = 0:1, ylim = 0:1)
abline(a = 0, b = 1, col = "darkblue")
```

The PE of cell lines cluster, but not super related (IMO) which is expected.
## Adjusting Growth Percentage for Plating Efficiency
```{r, cache = TRUE}
summary(CompleteTidy)
CompleteTidy <- CompleteTidy %>% mutate(R_2_1 = R_2_1 / W_2_1)
summary(CompleteTidy)
```

## Installed statmod to use ELDA/limdil functions

```{r}
library(statmod)
```

Important considerations:  
Observed = FALSE - indicates that we are dealing with a Poisson distribution, not a direct observation and confirmation of the plated cell number  

ELDA doesn't seem to be taking into account Plating Efficiency, so we might have to do that on our own, adjusting the "plated" number accordingly

**Going back to more raw dataset to undo some transformations**

```{r, cache = TRUE}
CompleteData <- CompleteData %>% mutate(LineID = paste(LineID, DaysPostInfection, sep = "_"))
head(CompleteData)
```

Need to calculate PE for each Line + Plate series and add to all rows

```{r}
WehiPE <- CompleteData %>% filter(Media == "WEHI")
head(WehiPE)
##THIS IS THE WRONG WAY TO CALCULATE PE!!! Doesn't assume Poisson distribution
#WehiPE <- WehiPE %>% mutate(PE = WellsGrown / TrueWellsPlated)
#head(WehiPE)

WehiELDA <- elda(response = WehiPE$WellsGrown, dose = WehiPE$CellsPerWell, tested = WehiPE$TrueWellsPlated, group = WehiPE$LineID, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
WehiELDAest <- as.data.frame(WehiELDA$CI)
head(WehiELDA)
WehiELDAest <- add_rownames(WehiELDAest, var = "LineID")
WehiELDAest <- WehiELDAest %>% mutate(LineID = as.factor(LineID))
#write.csv(WehiELDAest, "WehiELDAConfIntervals.csv")

#Drop Observations 11 and 12, incomplete due to contam
WehiELDAest <- WehiELDAest[c(-3,-4),]
head(WehiELDAest)
Odd <- seq(1,50,2)
Even <- seq(2,50,2)
PE2d <- WehiELDAest[Odd,3]
PE9d <- WehiELDAest[Even,3]
plot(PE2d$Estimate, PE9d$Estimate, xlab = "PE 2 Days post infection", ylab = "PE 9 Days post infection", xlim = c(0.8,3.5), ylim = c(0.8,3.5))
abline(a = 0, b = 1, col = "darkblue")

#Need to clean up the LineID field
WehiELDAest <- WehiELDAest %>% mutate(LineID = as.character(LineID))
WehiELDAest$LineID <- substr(WehiELDAest$LineID,7,nchar(WehiELDAest$LineID))


PEonly <- WehiELDAest %>% select(LineID, Estimate)
head(PEonly)


R10only <- CompleteData %>% filter(Media == "R10")
R10PE <- merge(R10only, PEonly, by="LineID")
head(R10PE)
```
Adjust Cells per Well by PE

```{r}
TidyData <- R10PE %>% mutate(PE_adjusted_cells = CellsPerWell / Estimate)
head(TidyData)
#write.csv(TidyData, file = "TidyData.csv")
```
Drop Samples 11 and 12 - Incomplete data due to plate contamination
```{r}
dim(TidyData)
TidyData <- TidyData %>% filter(!LineNum_legacy == 11) %>% filter(!LineNum_legacy == 12)
dim(TidyData)
```



Trial ELDA for 22_CSF2RB_R461C_2

```{r}
Trial <- TidyData %>% filter(LineNum_legacy == "22")
Trialout <- elda(Trial$WellsGrown, Trial$PE_adjusted_cells, Trial$TrueWellsPlated, Trial$LineID, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
plot(Trialout)
Trialout

#See if I can't clean up this code
Trialout2 <- elda(response = Trial$WellsGrown, dose = Trial$PE_adjusted_cells, tested = Trial$TrueWellsPlated, group = Trial$LineID, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
plot(Trialout2)

```

Massive ELDA

```{r}
MassELDA <- elda(response = TidyData$WellsGrown, dose = TidyData$PE_adjusted_cells, tested = TidyData$TrueWellsPlated, group = TidyData$LineID, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
plot(MassELDA)
ELDACI <- as.data.frame(MassELDA$CI)

head(ELDACI)
ELDACI <- add_rownames(ELDACI, var = "LineID")
ELDACI <- ELDACI %>% mutate(LineID = as.character(LineID))
#write.csv(MassELDA$CI, "ELDAConfIntervals.csv")
##These lines below aren't running, but I also don't need them, not sure if they ever worked
#TidyELDA <- merge(TidyData, ELDACI, by=1)
#head(TidyELDA)
```
Try to bin all runs together for ELDA
```{r}
TidyBin <- TidyData %>% mutate(Gene = paste(Transgene, Mutation, sep = "_"))
head(TidyBin)
BinELDA <- elda(response = TidyBin$WellsGrown, dose = TidyBin$PE_adjusted_cells, tested = TidyBin$TrueWellsPlated, group = TidyBin$Gene, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
BinCI <- as.data.frame(BinELDA$CI) %>% add_rownames(var = "LineID") %>% mutate(LineID = as.factor(LineID))
#write.csv(BinCI, "BinnedELDACI.csv")
```
Bin together technical replicates (day 2 and day 9 samples)
```{r}
TidyBin2 <- TidyData %>% mutate(Gene = paste(Transgene, Mutation, LineNum_legacy, sep = "_"))
head(TidyBin2)
Bin2ELDA <- elda(response = TidyBin2$WellsGrown, dose = TidyBin2$PE_adjusted_cells, tested = TidyBin2$TrueWellsPlated, group = TidyBin2$Gene, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
Bin2CI <- as.data.frame(Bin2ELDA$CI) %>% add_rownames(var = "LineID") %>% mutate(LineID = as.factor(LineID))
#write.csv(Bin2CI, "BinnedELDACI_3bins.csv")
```

