---
title: "Series78Analysis"
author: "KMWS"
date: "July 27, 2016"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(stringr)
library(statmod)

RawData <- read.csv("Series7_8_rawresults.csv")
summary(RawData)

RawComplete <- RawData %>% filter(!is.na(WellsGrown))
summary(RawComplete)
RawComplete <- RawComplete %>% mutate(LineDay = paste(LineNum_legacy, DaysPostInfection, sep = "_"))
```


```{r}
WEHIonly <- RawComplete %>% filter(Media == "WEHI")
summary(WEHIonly)
WEHIcomplete <- WEHIonly %>% filter(!is.na(WellsGrown))
Wpercents <- WEHIcomplete %>% transform(GrowthPercent = WellsGrown / WellsPlated)
hist(Wpercents$GrowthPercent)
```


```{r, cache = TRUE}
Wlowgrowth <- Wpercents %>% filter(GrowthPercent < 0.4)
Wlowgrowth[1,c(1,5,6,9,10)]
```
Comparing PE of cells plated at 2 days and 9 days post-infection

```{r, cache = TRUE}
Wpercents2 <- Wpercents %>% filter(DaysPostInfection == 2)
Wpercents9 <- Wpercents %>% filter(DaysPostInfection == 9)
summary(Wpercents2)
summary(Wpercents9)
hist(Wpercents2$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 2 days post-infection", breaks = 8, main = "WEHI 2 days post infection")
curve(dnorm(x, mean = mean(Wpercents2$GrowthPercent), sd = sd(Wpercents2$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)
hist(Wpercents9$GrowthPercent, xlim = c(0.2,1), xlab = "Percent Growth 9 days post-infection", breaks = 8, main = "WEHI 9 days post infection")
curve(dnorm(x, mean = mean(Wpercents9$GrowthPercent), sd = sd(Wpercents9$GrowthPercent)), add = TRUE, col = "darkblue", lwd = 2)

```

### These runs are not representative of each other

Need to calculate plating efficiency and adjust remaining counts accordingly

```{r}

WehiELDA <- elda(response = WEHIcomplete$WellsGrown, dose = WEHIcomplete$CellsPerWell, tested = WEHIcomplete$WellsPlated, group = WEHIcomplete$LineDay, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)

WehiELDAest <- as.data.frame(WehiELDA$CI)
head(WehiELDA)
WehiELDAest <- add_rownames(WehiELDAest, var = "LineDay")
WehiELDAest <- WehiELDAest %>% mutate(LineDay = as.factor(LineDay))
#write.csv(WehiELDAest, "WehiELDAConfIntervalsSeries78.csv")

#Need to clean up the LineID field
WehiELDAest <- WehiELDAest %>% mutate(LineDay = as.character(LineDay))
WehiELDAest$LineDay <- substr(WehiELDAest$LineDay,7,nchar(WehiELDAest$LineDay))

#Extract the PE estimate
PEonly <- WehiELDAest %>% select(LineDay, Estimate)
PEonly <- PEonly %>% rename(PlatingEff = Estimate)
head(PEonly)

#Merge with the R10 data
R10only <- RawComplete %>% filter(Media == "R10")
R10PE <- merge(R10only, PEonly, by="LineDay")
head(R10PE)

PEadjusted <- R10PE %>% mutate(TrueCellsPerWell = CellsPerWell / PlatingEff)
head(PEadjusted)


```

Calculate ELDA

```{r}

MassELDA <- elda(response = PEadjusted$WellsGrown, dose = PEadjusted$TrueCellsPerWell, tested = PEadjusted$WellsPlated, group = PEadjusted$LineDay, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
plot(MassELDA)
ELDACI <- as.data.frame(MassELDA$CI)

head(ELDACI)
ELDACI <- add_rownames(ELDACI, var = "LineDay")
ELDACI <- ELDACI %>% mutate(LineDay = as.character(LineDay))
#write.csv(MassELDA$CI, "ELDAConfIntervalsSeries78.csv")


```

The numbers are highly inflated and inconsistent with earlier experiments
I believe the concentration of IL-3 in the cells when plated might have been too high, resulting in excess IL-3 causing growth in too many 10,000 cell wells

```{r}
R10tenthousand <- R10only %>% filter(CellsPerWell == 10000)
R10tenthousand[,c(3,4,6,9,10)]

R10thousand <- R10only %>% filter(CellsPerWell == 1000)
R10thousand[,c(3,4,6,9,10)]
R10thousandDay2 <- R10thousand %>% filter(DaysPostInfection == 2)
R10thousandDay2[,c(3,4,6,9,10)]

```
Cut all 10000 cell wells from analysis
Cut 1,000 cell wells from Series 7 (Day 2)

```{r}
R10clean <- R10only %>% filter(!CellsPerWell == 10000)
head(R10clean)
Day2thousand <- R10clean$CellsPerWell == 1000 & R10clean$DaysPostInfection == 2
dim(R10clean)
R10cleaner <- R10clean[!Day2thousand,]
head(R10cleaner)
dim(R10cleaner)

#Merge with PE values
R10cleanerPE <- merge(R10cleaner, PEonly, by="LineDay")
head(R10cleanerPE)

R10cleanerPEadj <- R10cleanerPE %>% mutate(TrueCellsPerWell = CellsPerWell / PlatingEff)
head(R10cleanerPEadj)

```

Newer ELDA on limited data set

```{r}
MassELDAclean <- elda(response = R10cleanerPEadj$WellsGrown, dose = R10cleanerPEadj$TrueCellsPerWell, tested = R10cleanerPEadj$WellsPlated, group = R10cleanerPEadj$LineDay, observed = FALSE, confidence = 0.95, test.unit.slope = TRUE)
plot(MassELDAclean)
ELDACIclean <- as.data.frame(MassELDAclean$CI)

head(ELDACIclean)
ELDACIclean <- add_rownames(ELDACIclean, var = "LineDay")
ELDACIclean <- ELDACIclean %>% mutate(LineDay = as.character(LineDay))
#write.csv(MassELDAclean$CI, "ELDAConfIntervalsSeries78Censored.csv")


```

